---
title: "Homework2"
author: "Hao Zhou"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Homework2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 7,   # 调整图形宽度
  fig.height = 5,  # 调整图形高度
  fig.align = 'center',  # 居中对齐
  out.width = '80%',     # 输出宽度
  mar = c(4, 4, 2, 1)    # 调整图形边距 
)
```

## Question

### 3.4

The Rayleigh density [156, Ch. 18] is  
$$f(x)=\frac{x}{\sigma^2}e^{-x^2/(2\sigma^2)},\qquad x\geq0,\sigma>0.$$  
Develop an algorithm to generate random samples from a Rayleigh(σ) distribution. Generate Rayleigh(σ) samples for several choices of σ > 0 and check that the mode of the generated samples is close to the theoretical mode σ (check the histogram).  
Note that cdf $F(x)=1-e^{-x^2/(2\sigma^2)},x>0$.Then use the inverse transform  
$$u=1-e^{-x^2/(2\sigma^2)}\Longrightarrow F^{-1}(u)=\sigma(-2\log(1-u))^{1/2}.$$  
Obviously,$U$ and $1-U$ have the same distribution.

```{r}
n <- 3000
sigma <- seq(0.5,2,length = 4)
```

Samples of size n=3000 are generared for four choices of $\sigma$. Histograms below show the mode of the sample is
close to the theoretical mode $\sigma$.

```{r}
par(mfrow = c(2, 2))
for (i in 1:4){
  u <- runif(n)
  x <- sigma[i] * sqrt(-2*log(u))
  hist(x, breaks = "Scott", prob = TRUE, main = sigma[i])
  abline(v = sigma[i], lwd =1.5)
}
```

### 3.11

Generate a random sample of size 1000 from a normal location mixture. The components of the mixture have $N(0, 1)$ and $N(3, 1)$ distributions with mixing probabilities $p_1$ and $p_2 = 1 − p_1$. Graph the histogram of the sample with density superimposed, for $p_1 = 0.75$. Repeat with different values for $p_1$ and observe whether the empirical distribution of the mixture appears to be bimodal. Make a conjecture about the values of $p_1$ that produce bimodal mixtures.

```{r}
n <- 1000
p <- 0.75
mu <- sample(c(0, 3), size = 1000, replace =TRUE, prob = c(p, 1-p))
x <- rnorm(n, mu, 1)
```

Below is the histogram of the samole with density superimposed, for $p_1 = 0.75$.  

```{r}
hist(x, breaks = "Scott", prob = TRUE)
y <- sort(x)
fy <- p * dnorm(y) + (1-p) * dnorm(y, 3, 1)
lines(y, fy)
```

Repeat with different values for $p_1$.

```{r}
par(mfrow = c(3, 3))
p <- seq(0.1, 0.9, length = 9)
for (i in 1:9) {
  mu <- sample(c(0, 3),size = 1000, replace = TRUE, prob = c(p[i], 1-p[i]))
  x <- rnorm(n, mu, 1)
  hist(x, breaks = "Scott", prob = TRUE, main = paste("p=", p[i]))
  y <-sort(x)
  fy <- p[i] * dnorm(y) + (1-p[i]) * dnorm(y, 3, 1)
  lines(y, fy)
}
```

Through observing, we might conjecture that  $0.2 < p < 0.8$ produce bimodal mixtures.

### 3.20

A composed Poisson process is a stochastic process $\{X(t),t\ge 0\}$ that can be represented as the random sum $X(t)=\sum_{i=1}^{N(t)} Y_i,t\ge 0$, where $\{N(t),t\ge 0\}$ is a Poisson process and $Y_1, Y_2,\ldots$ are iid and independent of $\{N(t),t\ge 0\}$. Write a program to simulate a compound $Poisson(\lambda)-Gamma$ process (Y has a Gamma distribution). Estimate the mean and the variance of X(10) for several choices of the parameters and compare with the theoretical values.

```{r}
alpha <- 4
beta <- 2
lambda <- seq(2, 4, length = 3)
t <- 10
Pp <- function(lambda, t){
  tn <- rexp(1000, lambda)
  sn <- cumsum(tn)
  n <- min(which(sn > t))
  return(n-1)
}
x <- numeric(1000)
for (i in 1:3) {
  for (j in 1:1000) {
    N <- Pp(lambda[i], t)
    x[j] <- sum(rgamma(N, shape = alpha, rate = beta))
  }
  print(paste("Empirical value of the mean for lambda = ", lambda[i], "is", mean(x)))
  print(paste("Theoretical value of the mean for lambda = ", lambda[i], "is", lambda[i] * t * alpha/beta))
  print(paste("Empirical value of the variance for lambda = ", lambda[i], "is", var(x)))
  print(paste("Theoretical value of the varience for lambda = ", lambda[i], "is", lambda[i] * t * (alpha/beta^2 + (alpha/beta)^2)))
}
```


